Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2023-03-23 latex/comp_or.tex
Paper Comp OR update

@@

Parallelism is crucial for mobile computers to be capable of doing complicated operations while conserving energy. By leveraging parallelism, handheld computers can divide tasks into numerous smaller sections and execute these parts simultaneously, so utilizing multiple cores of the processor and offering a substantial performance improvement. This enables for more complicated applications and speedier performance, making mobile computers more competent than ever before.

@@

We also contributed by carrying out multiple experiments with the {\it Shims} heuristic on synthetic data based on real data from the {\it Brazilian Air Force} transportation history and a new procedure to minimize the distance from the next node destined pallet to the ramp door.

We converted {\it Shims} into a process-based parallel computing heuristic that quickly finds good solutions for a wide range of problem sizes. In addition, we applied a 3D packing procedure to guarantee that only the items that fit on each pallet are present in the solution.

@@

Air cargo transport involves several sub-problems that are difficult to solve. Recently, \cite{MesquitaSanches2023} modeled and solved the {\it Air Cargo Load Planning with Routing, Pickup and Delivery} (ACLP+RPDP) composed by four sub-problems: {\it Build-up Scheduling Problem}, {\it Air Palletization Problem} (APP), {\it Weight and Balance Problem} (WBP), and a special case of the {\it Traveling Salesman Problem}, similar to the proposed by \cite{kaspi2019}, where the profit per unit time is defined as the profit gained after completing the tour divided by the total time (cost in our case) required to complete the tour.

@@

As can be seen, so far \cite{LurkinSchyns2015} is the only work that simultaneously addresses an air cargo (WBP) and a flight itinerary (PDP) sub-problem. Although it is innovative, strong simplifications were imposed by the authors: in relation to loading, APP was ignored; with regard to routing, it is assumed a pre-defined flight plan restricted to two legs. It is important to note that these authors consider an aircraft with two doors, and the minimization of loading and unloading costs ({\bf MLU}) at the intermediate node was modelled through a container sequencing problem. Referring directly to this work, \cite[p. 409]{BrandtStefan2019} comment: {\it However, not even these sub-problems are acceptably solved for real-world problem sizes or the models omit some practically relevant constraints}.

@@

	\end{minipage}\hfill % these two lines must be close to each other
	\begin{minipage}{0.50\linewidth}
		\centering
		\includegraphics[scale=0.25]{Images/nodes.png}
		\captionof{figure}{A route between Brazilian airports}
		\label{fig:nodes}		
	\end{minipage}
\end{table}

@@

In this work, we raise four hypotheses for improvements to the work of \cite{MesquitaSanches2023}:
(1) Computer parallelism could improve {\it Shims}'s performance.
(2) The {\it Shims}' results may have been inferior to that; it could be that a sequential approach to pallet building may lead to a premature stop due to the torque constraints.
(3) The distances to move pallets on unloading in the next node could be minimized to benefit ground handling costs.
(4) A post-optimization procedure for 3D packing running in parallel mode would keep the running time adequate for operational use.

@@

\cite{Chan2006} carried out a case study with heterogeneous pallets. In order to minimise the total cost of shipping, they developed a 3D packing heuristic, with a loading plan for each pallet. Although the authors do not consider load balancing or positioning of pallets in the cargo hold, this method is relevant in commercial and industrial applications, where cargo items tend to be less dense.

@@

\cite{PaquaySchynsLimbourg2016} presented a mathematical modeling to optimize the loading of heterogeneous 3D boxes on pallets with a truncated parallelepipeds format. Its objective is to maximise the volume used in containers, considering load balancing constraints, the presence of fragile items and the possibility of rotating these boxes. \cite{PaquayLimbourgSchynsOliveira2018} developed some heuristics to solve this problem.

\cite{YangLiuGao2018} modelled the air transport problem as a 2D packing problem, and presented a heuristic for its optimization in several aircraft, considering load balancing in order to minimise fuel consumption.

@@

\cite{zhao2021} proposed a new modelling for WBP based on mixed integer programming. Instead of focusing on the center of gravity (CG) deviation, the authors consider the original CG envelope of the aircraft, with a linearization method for its non-linear constraints.

\cite{MesquitaSanches2023} contributed with a complex model and elaborate heuristics for the ACLP-RPDP that simultaneously solve 4 intractable sub-problems: APP, WBP, SPDP and TSP. They also compare the performances of four well known heuristics: {\it Ant Colony Optimization}, the {\it Noising Method Optimization}, the {\it Greedy Randomized Adaptive Search Procedure}, and {\it Tabu Search}. They also create a new heuristic called {\it Shims} which is fast and may be run in a handheld computer. They did not use computer parallelism to improve performance, nor tryed to minimize the distances from the next nodes destined pallets to the cargo ramp door.

As can be seen, except for \cite{MesquitaSanches2023}, the remaining works do not address air cargo palletization and load balancing with route optimization in a multi-leg flight plan, and none of them employ parallelization in their algorithms. This is the objective of our work: to reshape \cite{MesquitaSanches2023} solution process and algorithms to accommodate parallel features that can improve solution quality and performance and also include a new feature that is to minimize the distances between the next node's destined pallets and the cargo ramp door. We also intend to apply a 3D packing procedure to each pallet to guarantee that only the items that fit on each pallet are present in the solution.

@@

	\item We considered a unique pallet type: the {\it 463L Master Pallet}, a common size platform for bundling and moving air cargo. It is the primary air cargo pallet for more than 70 Air Forces and many air transport companies. This pallet has a capacity of $4500 kg$ and $14.8 m^3$, is equipped for locking into cargo aircraft rail systems, and includes tie-down rings to secure nets and cargo loads, which in total weighs $140 kg$. For more information, see {\tt www.463LPallet.com}.

	\item All items allocated on a pallet must have the same destination. A pallet which has not yet reached its destination may receive more items, although it is known that these operations of removing restraining nets increase handling time and the risk of improper delivery. We do not consider oversized cargo in this work, but only cargoes that fit on these pallets.

	\item Pallets with destinations set to the next node should be put as near as possible to the cargo ramp door.

	\item Finally, as we are interested in minimizing fuel costs while keeping the CG in its operational range, we disregarded some lower ones as handling costs.

@@

\section{The mathematical modelling}
\label{sec4}

Given the assumptions, scenarios and parameters described in the previous section, we are ready to present the mathematical modelling of ACLP+RPDP.

@@

Let $S_K = \{s: \{1, \dots, K\} \rightarrow \{1, \dots, K\} \}$ be the set of $K!$ permutations, which correspond to all possible tours (or itineraries) that have $l_0$ as origin and end, passing through the other $K$ nodes. {\color{blue} $k$ is the node index and $\pi(k)$ is the node in position $k$ of tour $\pi$}.

Let $M = \{p_1, p_2, \ldots, p_m \}$ the set of $m$ pallets. Each pallet $p_i$, $1 \leq i \leq m$, has weight capacity $p_i.w$, volume capacity $p_i.v$, pallet destinations $p_i.to[{\pi(k)}]$, $0 \leq k \leq K$, and distance to the CG of aircraft $p_i.d$. $p_i.to[{\pi(k)}]$ denotes that pallet $p_i$ may assume a different destination in each node ${\pi(k)}$.

Let $N_{\pi(k)} = \{t^{\pi(k)}_1, t^{\pi(k)}_2, \ldots, t^{\pi(k)}_{n_{\pi(k)}} \}$ be the set of $n_{\pi(k)}$ items to be loaded in node $l_{\pi(k)}$, $0 \leq k \leq K$. Each item $t^{\pi(k)}_j$, $1 \leq j \leq n_{\pi(k)}$, has score $t^{\pi(k)}_j.s$, weight $t^{\pi(k)}_j.w$, volume $t^{\pi(k)}_j.v$, and destination $t^{\pi(k)}_j.to \in L_{\pi(k)}$. Let $N = \bigcup_{0 \leq k \leq K} N_{\pi(k)}$ be the set of items of all nodes along a tour.

Let $Q_{\pi(k)} = \{a^{\pi(k)}_1, a^{\pi(k)}_2, \ldots, a^{\pi(k)}_{m_{\pi(k)}} \}$ be the set of consolidated items loaded in $m_k \leq m$ pallets when the aircraft arrives at node $l_{\pi(k)}$, with $0 \leq k \leq K$. $a^{\pi(k)}_i$, $1 \leq i \leq m_{\pi(k)}$, is the group of picked-up items that were allocated on pallet $p_i$ in some of the previous nodes. $a^{\pi(k)}_i$ has total weight $a^{\pi(k)}_i.w$, total volume $a^{\pi(k)}_i.v$, and destination $a^{\pi(k)}_i.to \in L_{\pi(k)} \cup \{l_{\pi(k)}\}$. If $a^{\pi(k)}_i.to = l_{\pi(k)}$, then $a^{\pi(k)}_i$ is unloaded, and $p_i$ will be available for reloading; otherwise, $a^{\pi(k)}_i$ remains on the aircraft, eventually in another pallet, and with items of $N_{\pi(k)}$ having the same destination.

Let $X_{ij}^{\pi(k)}$ and $Y_{iq}^{\pi(k)}$ be binary variables, where $0 \leq k \leq K$, $1 \leq j \leq n_{\pi(k)}$, $1 \leq i \leq m$ and $1 \leq q \leq m_{\pi(k)}$. $X_{ij}^{\pi(k)} = 1$ if $t_j^{\pi(k)}$ is assigned to $p_i$ in node $l_{\pi(k)}$, and 0 otherwise. $Y_{iq}^{\pi(k)} = 1$ if $a_q^{\pi(k)}$ is assigned to $p_i$ in node $l_{\pi(k)}$, and 0 otherwise. By definition, $Y_{iq}^0=0$. Allocations of items or consolidated items to pallets in node $l_{\pi(k)}$ can be seen as a bipartite graph $G_{\pi(k)}(V_{\pi(k)}, E_{\pi(k)})$, where $V_{\pi(k)} = M \cup N_{\pi(k)} \cup Q_{\pi(k)}$, $E_{\pi(k)} = E^N_{\pi(k)} \cup E^Q_{\pi(k)}$, $(p_i, t_j^{\pi(k)}) \in E^N_{\pi(k)}$ if $X_{ij}^{\pi(k)} = 1$, and $(p_i, a_q^{\pi(k)}) \in E^Q_{\pi(k)}$ if $Y_{iq}^{\pi(k)} = 1$.

@@

\begin{equation} \label{eq:scores}
	\tilde{s} = \sum_{k=0}^{K} \sum_{i=1}^{m} \sum_{j=1}^{n_{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.s
\end{equation}

\begin{equation} \label{eq:costs}
	\tilde{c} = c_{0,\pi(1)}\times(1+c_g\times|\epsilon_0|) + \sum_{k=1}^{K-1} [ c_{\pi(k), \pi(k+1)}\times(1+c_g\times|\epsilon_{\pi(k)}|) ] + c_{\pi(K),0}\times(1+c_g\times|\epsilon_K|)
\end{equation}

@@

\begin{equation} \label{eq:tau}
\tau_{\pi(k)} = \sum_{i=1}^{m}[ p_i.d \times (\sum_{j=1}^{n_{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w +  \sum_{q=1}^{m} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w)];\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:eps}
\epsilon_{\pi(k)} = \frac{\tau_{\pi(k)}}{maxW \times limit^{CG}_{long}};\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp11}
	L_0 = L; L_{\pi(k)} = L_{{\pi(k-1)}} - \{l_{\pi(k)}\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

@@

\begin{equation} \label{eq:pdp12}
	X_{ij}^{\pi(k)} = 0 \ \mbox{if} \ t_j^{\pi(k)}.to \notin L_{\pi(k)}; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n_{\pi(k)}\}; \ k \in \{1,2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp9}
	Y_{iq}^{\pi(k)} = 0 \ \mbox{if} \ a_i^{\pi(k)}.to \notin L_{\pi(k)}; \ i \in \{1, 2, \ldots, m\}; \ q \in \{1, 2, \ldots, m_{\pi(k)}\}; \ k \in \{ 1, 2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:cons2}
	a_i^{\pi(k+1)}.w = \sum_{j=1}^{n_{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w + \sum_{q=1}^{m_{\pi(k)}} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w;  \ i \in \{1, 2, \ldots, m_{\pi(k)}\}; k \in \{0, 1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:cons3}
	a_i^{\pi(k+1)}.v = \sum_{j=1}^{n_{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.v + \sum_{q=1}^{m_k} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.v;  \ i \in \{1, 2, \ldots, m_{\pi(k)}\}; k \in \{0, 1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:cons5}
	a_i^{\pi(k+1)}.to = t^{\pi(k)}_j.to  \ \mbox{if} \ X^{\pi(k)}_{ij} = 1  \ \mbox{and} \ t^{\pi(k)}_j.to \in L_{\pi(k+1)};  \ i \in \{1, 2, \ldots, m_{\pi(k)}\}; \ j \in \{1, 2, \ldots, n_{\pi(k)}\}; \ k \in \{0,1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:LatIt}
	LatIt_{\pi(k)} = \sum_{i=1}^{m} \sum_{j=1}^{n_{\pi(k)}} ( X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w \times (i\%2) - X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w \times (i+1)\%2 ); \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:LatCons}
	LatCons_{\pi(k)} =  \sum_{i=1}^{m} \sum_{q=1}^{m_{\pi(k)}}  ( Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w \times (i\%2) - Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w \times (i+1)\%2); \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:torqlat}
	s.t.: d^{CG}_{pallet} \times | LatIt_{\pi(k)} + LatCons_{\pi(k)} | \leq  \sum_{i=1}^{m}p_i.w \times limit^{CG}_{lat}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:torqlong}
	s.t.: |\tau_{\pi(k)}| \leq maxW \times limit^{CG}_{long};\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:payload}
	s.t.: \sum_{i=1}^{m} (\sum_{j=1}^{n_{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w + \sum_{q=1}^{m_k} Y_{iq}^{\pi(k)} \times a_q^k.w ) \leq maxW; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app2}
	s.t.: \sum_{j=1}^{n_{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w + \sum_{q=1}^{m_k} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w  \leq p_i.w; \ i \in \{1, 2, \ldots, m_{\pi(k)}\}; \ {\pi(k)} \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app3}
	s.t.: \sum_{j=1}^{n_{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.v + \sum_{q=1}^{m_{\pi(k)}} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.v  \leq\ p_i.v; \ i \in \{1, 2, \ldots, m_{\pi(k)}\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app4}
	s.t.: \sum_{i=1}^{m} X_{ij}^{\pi(k)} \leq 1; \ j \in \{1, 2, \ldots, n_{\pi(k)}\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app5}
	s.t.:  Y_{iq}^{\pi(k)} = 1 \ \mbox{if} \ a^{\pi(k)}_q.to \in L_{\pi(k)}; \ q \in \{1, 2, \ldots, m_{\pi(k)}\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp8}
	s.t.: p_i.to[{\pi(k)}] = t^{\pi(k)}_j.to\ \mbox{if} \ X_{ij}^{\pi(k)} = 1; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n_{\pi(k)}\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp2}
	s.t.:  p_i.to[{\pi(k)}] = a^{\pi(k)}_q.to\ \mbox{if} \ Y_{iq}^{\pi(k)} = 1; \ i \in \{1, 2, \ldots, m\};\ q \in \{1, 2, \ldots, m_{\pi(k)}\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

>>>>

\usepackage{siunitx}

@@

Parallelism is crucial for mobile computers to be capable of doing complicated operations while conserving energy. By leveraging parallelism, handheld computers can divide tasks into numerous smaller sections and execute these parts simultaneously, so utilizing multiple cores of the processor and offering a substantial performance improvement. This enables for more complicated applications and speedier execution, making mobile computers more competent than ever before.

@@

We also contributed by carrying out multiple experiments with the multi-processing heuristics on synthetic data based on real data from the {\it Brazilian Air Force} transportation history and a new procedure to minimize the distance from the next node destined pallet to the ramp door.

We converted {\it Shims} into a process-based parallel computing heuristic that quickly finds good solutions for a wide range of problem sizes. We also solved a three-dimensional packing problem in the item-pallet allocations yielded by {\it Shims}, which can guarantee that the palletization plan will include sets of items that actually fit into the pallets.

@@

Air cargo transport involves several sub-problems that are difficult to solve. Recently, \cite{MesquitaSanches2023} modeled and solved the {\it Air Cargo Load Planning with Routing, Pickup and Delivery} (ACLP+RPDP) composed by four sub-problems: {\it Build-up Scheduling Problem}, {\it Air Palletization Problem} (APP), {\it Weight and Balance Problem} (WBP), and a special case of the {\it Traveling Salesman Problem}, similar to the proposed by \cite{kaspi2019}, where the profit per unit time is defined as the gain after completing the tour divided by the total time (cost in our case) required to complete the tour.

@@

As can be seen, so far \cite{LurkinSchyns2015} is the only work that simultaneously addresses an air cargo (WBP) and a flight itinerary (PDP) sub-problem. Although it is innovative, strong simplifications were imposed by the authors: in relation to loading, APP was ignored; with regard to routing, it is assumed a pre-defined flight plan restricted to two legs. It is important to note that these authors consider an aircraft with two doors, and the minimization of loading and unloading costs ({\bf MLU}) at the intermediate node was modeled through a container sequencing problem. Referring directly to this work, \cite[p. 409]{BrandtStefan2019} comment: {\it However, not even these sub-problems are acceptably solved for real-world problem sizes or the models omit some practically relevant constraints}.

@@

	\end{minipage}\hfill % these two lines must be close to each other
	\begin{minipage}{0.50\linewidth}
		\centering
		\includegraphics[scale=0.25]{Images/nodes.png}
		\captionof{figure}{A route between Brazilian airports}
		\label{fig:nodes}		
	\end{minipage}
\end{table}
\label{hypotheses}

@@

In this work, we raise four hypotheses for improvements to the work of \cite{MesquitaSanches2023}: 
(1) Computer parallelism could improve {\it Shims}'s performance.
(2) The {\it Shims}' results may have been inferior to that; it could be that a sequential approach to pallet building may lead to a premature stop due to the torque constraints.
(3) A procedure for 3D packing running in parallel mode would keep the run time adequate for operational use.
(4) The distances to move pallets on unloading in the next node could be minimized to benefit ground handling costs.

@@

\cite{Chan2006} carried out a case study with heterogeneous pallets. In order to minimize the total cost of shipping, they developed a 3D packing heuristic, with a loading plan for each pallet. Although the authors do not consider load balancing or positioning of pallets in the cargo hold, this method is relevant in commercial and industrial applications, where cargo items tend to be less dense.

@@

\cite{PaquaySchynsLimbourg2016} presented a mathematical modeling to optimize the loading of heterogeneous 3D boxes on pallets with a truncated parallelepipeds format. Its objective is to maximize the volume used in containers, considering load balancing constraints, the presence of fragile items and the possibility of rotating these boxes. \cite{PaquayLimbourgSchynsOliveira2018} developed some heuristics to solve this problem.

\cite{YangLiuGao2018} modeled the air transport problem as a 2D packing problem, and presented a heuristic for its optimization in several aircraft, considering load balancing in order to minimize fuel consumption.

@@

\cite{zhao2021} proposed a new model for WBP based on mixed integer programming. Instead of focusing on the center of gravity (CG) deviation, the authors consider the original CG envelope of the aircraft, with a linearization method for its non-linear constraints.

\cite{MesquitaSanches2023} contributed with a complex model and elaborate heuristics for the ACLP-RPDP that simultaneously solve 4 intractable sub-problems: APP, WBP, SPDP and TSP. They also compare the performances of four well known heuristics: {\it Ant Colony Optimization}, the {\it Noising Method Optimization}, the {\it Greedy Randomized Adaptive Search Procedure}, and {\it Tabu Search}. They also create a new heuristic called {\it Shims} which is fast and may be run in a handheld computer. They did not use multi-core parallelism to improve performance, nor tried to minimize the distances from the next nodes destined pallets to the cargo ramp door, nor tried to perform a 3D packing to keep in solution only items that fit in the pallets usable volume.

As can be seen, except for \cite{MesquitaSanches2023}, the remaining works do not address air cargo palletization and load balancing with route optimization in a multi-leg flight plan, and none of them employ parallelization in their algorithms. This is the objective of our work: to reshape \cite{MesquitaSanches2023} solution process and algorithms to accommodate parallel features that can improve solution quality and performance, include a new feature that is to minimize the distances between the next node's destined pallets and the cargo ramp door, and also include a 3D packing procedure to guarantee that only items that fit into the pallets usable volume are planned to palletization.

@@

	\item We considered a unique pallet type: the {\it 463L Master Pallet}, a common size platform for bundling and moving air cargo. It is the primary air cargo pallet for more than 70 Air Forces and many air transport companies. This pallet has a capacity of $4500 kg$ and $14.8 m^3$ (depending on the aircraft ceiling height), is equipped for locking into cargo aircraft rail systems, and includes tie-down rings to secure nets and cargo loads, which in total weighs $140 kg$. For more information, see {\tt www.463LPallet.com}.

	\item All items allocated on a pallet must have the same destination. A pallet which has not yet reached its destination may receive more items, although it is known that these operations of removing restraining nets increase handling time and the risk of improper delivery. We do not consider over-sized cargo in this work, but only cargoes that fit on these pallets.

@@

\section{The mathematical modeling}
\label{sec4}

Given the assumptions, scenarios and parameters described in the previous section, we are ready to present the mathematical modeling of ACLP+RPDP.

@@

Let $S_K = \{s: \{1, \dots, K\} \rightarrow \{1, \dots, K\} \}$ be the set of $K!$ permutations {\color{blue} indexed by $\pi$}, which correspond to all possible tours (or itineraries) that have $l_0$ as origin and end, passing through the other $K$ nodes.

{\color{blue} $k$ is the node index and $\pi(k)$ is the node in position $k$ of tour $\pi$. $\tau^{\pi(k)}$ is the total torque applied by the loaded pallets to the aircraft in this node. }

Let $M^{\pi(k)} = \{p_1, p_2, \ldots, p_m \}$ the set of $m$ pallets. Each pallet $p_i$, $1 \leq i \leq m$, has weight capacity $p_i.w$, volume capacity $p_i.v$, pallet destinations $p_i.to^{\pi(k)}$, $0 \leq k \leq K$, and distance to the CG of aircraft $p_i.d$. $p_i.to^{\pi(k)}$ denotes that pallet $p_i$ may assume a different destination in each node ${\pi(k)}$.

Let $N^{\pi(k)} = \{t^{\pi(k)}_1, t^{\pi(k)}_2, \ldots, t^{\pi(k)}_{n^{\pi(k)}} \}$ be the set of $n^{\pi(k)}$ items to be loaded in node $l^{\pi(k)}$, $0 \leq k \leq K$. Each item $t^{\pi(k)}_j$, $1 \leq j \leq n^{\pi(k)}$, has score $t^{\pi(k)}_j.s$, weight $t^{\pi(k)}_j.w$, volume $t^{\pi(k)}_j.v$, and destination $t^{\pi(k)}_j.to \in L^{\pi(k)}$. Let $N = \bigcup_{0 \leq k \leq K} N^{\pi(k)}$ be the set of items of all nodes along a tour.

Let $Q^{\pi(k)} = \{a^{\pi(k)}_1, a^{\pi(k)}_2, \ldots, a^{\pi(k)}_{m} \}$ be the set of consolidated items loaded in $m_k \leq m$ pallets when the aircraft arrives at node $l^{\pi(k)}$, with $0 \leq k \leq K$. $a^{\pi(k)}_i$, $1 \leq i \leq m$, is the group of picked-up items that were allocated on pallet $p_i$ in some of the previous nodes. $a^{\pi(k)}_i$ has total weight $a^{\pi(k)}_i.w$, total volume $a^{\pi(k)}_i.v$, and destination $a^{\pi(k)}_i.to \in L^{\pi(k)} \cup \{l^{\pi(k)}\}$. If $a^{\pi(k)}_i.to = l^{\pi(k)}$, then $a^{\pi(k)}_i$ is unloaded, and $p_i$ will be available for reloading; otherwise, $a^{\pi(k)}_i$ remains on the aircraft, eventually in another pallet, and with items of $N^{\pi(k)}$ having the same destination.

Let $X_{ij}^{\pi(k)}$ and $Y_{iq}^{\pi(k)}$ be binary variables, where $0 \leq k \leq K$, $1 \leq j \leq n^{\pi(k)}$, $1 \leq i \leq m$ and $1 \leq q \leq m$. $X_{ij}^{\pi(k)} = 1$ if $t_j^{\pi(k)}$ is assigned to $p_i$ in node $l^{\pi(k)}$, and 0 otherwise.

$Y_{iq}^{\pi(k)} = 1$ if $a_q^{\pi(k)}$ is assigned to $p_i$ in node $l^{\pi(k)}$, and 0 otherwise. By definition, $Y_{iq}^0=0$.

Allocations of items or consolidated to pallets in node $l^{\pi(k)}$ can be seen as a bipartite graph $G^{\pi(k)}(V^{\pi(k)}, E^{\pi(k)})$,
where $V^{\pi(k)} = M^{\pi(k)} \cup N^{\pi(k)} \cup Q^{\pi(k)}$, $E^{\pi(k)} = E_N^{\pi(k)} \cup E_Q^{\pi(k)}$,
$(p_i, t_j^{\pi(k)}) \in E_N^{\pi(k)}$
if $X_{ij}^{\pi(k)} = 1$, and $(p_i, a_q^{\pi(k)}) \in E_Q^{\pi(k)}$ if $Y_{iq}^{\pi(k)} = 1$.

@@

\begin{equation} \label{eq:scores}
	\tilde{s} = \sum_{k=0}^{K} \sum_{i=1}^{m} \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.s
\end{equation}

\begin{equation} \label{eq:costs}
	\tilde{c} = c_{0,\pi(1)}\times(1+c_g\times|\epsilon_0|) + \sum_{k=1}^{K-1} [ c^{\pi(k), \pi(k+1)}\times(1+c_g\times|\epsilon^{\pi(k)}|) ] + c^{\pi(K),0}\times(1+c_g\times|\epsilon_K|)
\end{equation}

@@

\begin{equation} \label{eq:tau}
\tau^{\pi(k)} = \sum_{i=1}^{m}[ p_i.d \times (\sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w +  \sum_{q=1}^{m} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w)];\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:eps}
\epsilon^{\pi(k)} = \frac{\tau^{\pi(k)}}{maxW \times limit^{CG}_{long}};\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp11}
	L_0 = L; L^{\pi(k)} = L^{\pi(k-1)} - \{l^{\pi(k)}\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

@@

\begin{equation} \label{eq:pdp12}
	X_{ij}^{\pi(k)} = 0 \ \mbox{if} \ t_j^{\pi(k)}.to \notin L^{\pi(k)}; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n^{\pi(k)}\}; \ k \in \{1,2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp9}
	Y_{iq}^{\pi(k)} = 0 \ \mbox{if} \ a_i^{\pi(k)}.to \notin L^{\pi(k)}; \ i \in \{1, 2, \ldots, m\}; \ q \in \{1, 2, \ldots, m\}; \ k \in \{ 1, 2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:cons2}
	a_i^{\pi(k+1)}.w = \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w + \sum_{q=1}^{m} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w;  \ i \in \{1, 2, \ldots, m\}; k \in \{0, 1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:cons3}
	a_i^{\pi(k+1)}.v = \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.v + \sum_{q=1}^{m_k} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.v;  \ i \in \{1, 2, \ldots, m\}; k \in \{0, 1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:cons5}
	a_i^{\pi(k+1)}.to = t^{\pi(k)}_j.to  \ \mbox{if} \ X^{\pi(k)}_{ij} = 1  \ \mbox{and} \ t^{\pi(k)}_j.to \in L_{\pi(k+1)};  \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n^{\pi(k)}\}; \ k \in \{0,1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:LatIt}
	LatIt^{\pi(k)} = \sum_{i=1}^{m} \sum_{j=1}^{n^{\pi(k)}} ( X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w \times (i\%2) - X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w \times (i+1)\%2 ); \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:LatCons}
	LatCons^{\pi(k)} =  \sum_{i=1}^{m} \sum_{q=1}^{m}  ( Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w \times (i\%2) - Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w \times (i+1)\%2); \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:torqlat}
	s.t.: d^{CG}_{pallet} \times | LatIt^{\pi(k)} + LatCons^{\pi(k)} | \leq  \sum_{i=1}^{m}p_i.w \times limit^{CG}_{lat}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:torqlong}
	s.t.: |\tau^{\pi(k)}| \leq maxW \times limit^{CG}_{long};\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:payload}
	s.t.: \sum_{i=1}^{m} (\sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w + \sum_{q=1}^{m_k} Y_{iq}^{\pi(k)} \times a_q^k.w ) \leq maxW; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app2}
	s.t.: \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w + \sum_{q=1}^{m_k} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w  \leq p_i.w; \ i \in \{1, 2, \ldots, m\}; \ {\pi(k)} \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app3}
	s.t.: \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.v + \sum_{q=1}^{m} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.v  \leq\ p_i.v; \ i \in \{1, 2, \ldots, m\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app4}
	s.t.: \sum_{i=1}^{m} X_{ij}^{\pi(k)} \leq 1; \ j \in \{1, 2, \ldots, n^{\pi(k)}\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app5}
	s.t.:  Y_{iq}^{\pi(k)} = 1 \ \mbox{if} \ a^{\pi(k)}_q.to \in L^{\pi(k)}; \ q \in \{1, 2, \ldots, m\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp8}
	s.t.: p_i.to^{\pi(k)} = t^{\pi(k)}_j.to\ \mbox{if} \ X_{ij}^{\pi(k)} = 1; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n^{\pi(k)}\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp2}
	s.t.:  p_i.to^{\pi(k)} = a^{\pi(k)}_q.to\ \mbox{if} \ Y_{iq}^{\pi(k)} = 1; \ i \in \{1, 2, \ldots, m\};\ q \in \{1, 2, \ldots, m\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}
